Growing up with Nell: A Narrative Interface for Literacy

### 2012-03-11 paper.tex
More tweaks for space.  References now take up only 1 col (12.5% total space).

\subsection{Narrative}

Experience with our previous educational software
system~\cite{flores:uruguay,idb:peru} has shown gratifying uptake and
use by children.  But we are often disappointed that some of our
favorite pedagogical tools included in the systems were not being
widely used.  Enthusiastic teachers would spark intensive use in
certain classrooms, but children had difficulty finding their way into
the material without strong guidance.

@@

Nell's overall story is a multi-character serial adventure, taking
cues from \textit{The Diamond Age} and
from the UNIVERSE narrative-generation system \cite{lebowitz:universe84}.
Each of the several characters represents a specific skill or subject
area, and each adventure represents about a year's curriculum.  This
design also improves modularity and allows updates and decentralized
authorship.

\begin{figure}
\centering
\epsfig{file=roger1.eps, height=1.5in} % width=3.2in
\caption{Illustration of Nell's primary interface.}\label{fig:nell}
\end{figure}

@@

\textit{Avatar Services} stores all choices made by the child and make them
accessible to story modules and activities.  Accomplishments
(completed lessons) are also stored.  Nell can thus provide multiple
alternative story modules for a given plot point/lesson, and select
between them based on either the explicit choices made by the child
(``fractions in outer space,'' if the child previously indicated a
preference for space-themed lessons) or prior success with a given
lesson style (a large number of accomplishments in musical-rhythmic
tasks suggests a rhythmic approach to fractions).  In this manner we
can engage the child's interests and cater to different learning
styles~\cite{gardner:mi}.

@@

We leverage AIML~\cite{aiml:2005},
the markup language developed for the three-time Loebner Prize-winning
AliceBot.%
\footnote{Wilcox identifies several shortcomings with the expressive
  power of AIML, proposing ChatScript as a
  replacement~\cite{wilcox:2010}.  At present AIML's community and
  diversity of third-party implementations make it preferable.}
Each story module can include AIML fragments extending the
conversational capabilities of Nell's agents.

@@

The TinkRBook system~\cite{chang:tinkrbook} shares a similar focus on
personalization and literacy.  Project Nell attempts to scale these
ideas with a modular extensible narrative system and a broader range
of pedagogic material.

>>>>

\subsection{Narrative}
%used to \cite{flores:uruguay,idb:peru} below; removed for space

Experience with our previous educational software
system has shown gratifying uptake and
use by children.  But we are often disappointed that some of our
favorite pedagogical tools included in the systems were not being
widely used.  Enthusiastic teachers would spark intensive use in
certain classrooms, but children had difficulty finding their way into
the material without strong guidance.

@@

Nell's overall story is a multi-character serial adventure, taking
cues from \textit{The Diamond Age} and
from the UNIVERSE narrative-generation system \cite{lebowitz:universe85}.
Each of the several characters represents a specific skill or subject
area, and each adventure represents about a year's curriculum.  This
design also improves modularity and allows updates and decentralized
authorship.

\begin{figure}
\centering
\epsfig{file=roger1.eps, height=1.5in} % width=3.2in
\caption{Diagram of Nell's primary interface.}\label{fig:nell}
\end{figure}

@@

\textit{Avatar Services} stores all choices made by the child and make them
accessible to story modules and activities.  Accomplishments
(completed lessons) are also stored.  Nell can thus provide multiple
alternative story modules for a given plot point/lesson, and select
between them based on either the explicit choices made by the child
(``fractions in outer space,'' if the child previously indicated a
preference for space-themed lessons) or prior success with a given
lesson style (a large number of accomplishments in musical-rhythmic
tasks suggests a rhythmic approach to fractions).  In this manner we
can engage the child's interests and cater to different learning
styles.%~\cite{gardner:mi}.%%%%%%% <- lead with this.

@@

We leverage AIML~\cite{aiml:2005},
the markup language developed for the three-time Loebner Prize-winning
AliceBot.%
\footnote{Wilcox identifies several shortcomings with the expressive
  power of AIML, proposing ChatScript as a
  replacement~\cite{wilcox:2010}.  AIML's community and
  diversity of third-party implementations make it preferable at present.}
Each story module can include AIML fragments extending the
conversational capabilities of Nell's agents.

@@

The TinkRBook system~\cite{chang:tinkrbook} shares a similar focus on
personalization and literacy.  Nell attempts to scale these
ideas with a modular extensible narrative system and a broader range
of pedagogic material.
